A performance assessment of evolutionary algorithms in volunteer computing environments: the importance of entropy

### 2016-11-14 EvoPar2017/paper.Rnw
Adds entropy linear model

%
%
\begin{figure}[htbp]
\centering
<<zipf, cache=FALSE,echo=FALSE,fig.height=4>>=
ggplot(ips.data,aes(x=idx,y=PUTs,group=Experiment,color=Experiment))+geom_point()+scale_y_log10()+ scale_colour_discrete(labels=c('base','graph1','graph2','reboot','cachereboot'))+scale_x_log10()
@ %def graph for IPs and puts
%
\caption{Log-log plot of ranked and normalized number of contributions (PUTs) per IP.\label{fig:zipf}}
\end{figure}
%
Every run had slightly different conditions, although in some cases it
was just a server reboot and a new round of announcements. Each run
batch included at least 30 completions, that is, evolution until the
solution was found. The number of runs was controlled by polling an
URL that indicated the number of runs made in the present batch so
far. Every volunteer contributed a certain number of individuals
(represented by the PUT HTTP request) to the server. The ranked number
of contributions per unique IP is shown in Figure \ref{fig:zipf} in a
log-log scale. Besides following the usual Zipf's law found in our
previous work and thus having a very similar appearance for all 5 sets, it
is interesting to see how some sets have a number of contributions per
IP that are slightly superior to the rest, specially so in {\tt
  graph2}. The last two sets, {\tt reboot} and {\tt cachereboot}, also
show a regime in the middle of the graph that shows that some users
are contributing more than in other similar sets. In general, these
differences in behavior are due more to the specific users that show
up in these experiments than to general laws, but it should also be
noted that those behaviors can be encouraged by design.

@@

In order to have enough data to perform more analysis on it, a new
experiment was performed along several days with, eventually, several
hundreds of persons participating on it. It was announced by several
popular Twitter figures and found its way into Reddit. The number of
PUTs per minute are shown in \ref{fig:minute}. Sudden peaks in them
are due to new announcements. Please check also the dot color, which
corresponds to the {\em ratio} or, as indicated above, the actual
number of contributing PUTs, those that have not been discarded due to
repeated sending by clients or (briefly) belonging to the previous
instance of the experiment. It reveals that, in some cases where the
amount of PUTs really increases, they are actually not contributing so
much, while light dots, corresponding to a ration close to 1, usually
stay in the lowest part of the graphic.

This is interesting in the sense that in indicates that there might be
something more to performance that just the number of volunteers
contributing. In this experiment we include an additional measure:
user-contribution entropy, which is measure by computing the
compression entropy of the user contributions every for the whole
experiment.  This is computed converting the sequence of the number of
PUTs into a string and then compressing it using {\tt Zlib}, a
standard compression library. Since the compression rate is related to
the randomness of the sequence, a higher compression rate will
correspond to an experiment with a low entropy, while a lower
compression rate will indicate that there has been a certain amount of
unexpected changes during the experiment. Although less popular than
Jenssen-Shannon entropy, this way of measuring it is widely used
\cite{ornstein1993entropy} in case where we want to compute the
entropy of sequences of events, such as arrhythmias
\cite{baumert2004forecasting}. Unlike Jenssen-Shannon entropy, the
range of the compression entropy will vary between 0 (lower, or worse
compression rate) and 1 (no compression). 
%
\begin{figure}[ht!b]
\centering
<<entropy, cache=FALSE,echo=FALSE>>=
max <- max(new.experiment.data$IPs)
exp.palette <- colorRampPalette(brewer.pal(12,"Paired"))
ggplot(data=new.experiment.data)+geom_point(aes(x=entropy,y=milliseconds,size=PUTs,color=IPs))+scale_x_log10()+scale_y_log10()+scale_color_gradient(limits=c(1,max),low='red')
@ %def entropy data
%
\caption{Entropy of the user contribution vs. time to solution in the
  volunteer computing experiment. \label{fig:entropy}}
\end{figure}

>>>>

%
%
\begin{figure}[htbp]
\centering
<<zipf, cache=FALSE,echo=FALSE,fig.height=6>>=
ggplot(ips.data,aes(x=idx,y=PUTs,group=Experiment,color=Experiment))+geom_point()+scale_y_log10()+ scale_colour_discrete(labels=c('base','graph1','graph2','reboot','cachereboot'))+scale_x_log10()
@ %def graph for IPs and puts
%
\caption{Log-log plot of ranked and normalized number of contributions (PUTs) per IP.\label{fig:zipf}}
\end{figure}
%
Every run had slightly different conditions, although in some cases it
was just a server reboot and a new round of announcements. Each run
batch included at least 30 completions, that is, evolution until the
solution was found. The number of runs was controlled by polling an
URL that indicated the number of runs made in the present batch so
far. Every volunteer contributed a certain number of individuals
(represented by the PUT HTTP request) to the server. The ranked number
of contributions per unique IP is shown in Figure \ref{fig:zipf} in a
log-log scale. Besides following the usual Zipf's law found in our
previous work and thus having a very similar appearance for all 5 sets, it
is interesting to see how some sets have a number of contributions per
IP that are slightly superior to the rest, specially so in {\tt
  graph2}. The last two sets, {\tt reboot} and {\tt cachereboot}, also
show a regime in the middle of the graph that shows that some users
are contributing more than in other similar sets. In general, these
differences in behavior are due more to the specific users that show
up in these experiments than to general laws, but it should also be
noted that those behaviors can be encouraged by design.

@@

In order to have enough data to perform more analysis on it, a new
experiment was performed along several days with, eventually, several
hundreds of persons participating on it. It was announced by several
popular Twitter figures and found its way into Reddit. The number of
PUTs per minute are shown in \ref{fig:minute}. Sudden peaks in them
are due to new announcements in social networks, usually; this is one
of the problems with this kind of volunteer computing systems: they
feed on the continuous influx of new users; however, this can be
automatized and also encouraged, and, as shown, once enough users have
participated it is sometimes easier to find new ones.

In Figure \ref{fig:minute} check
also the dot color, which 
corresponds to the {\em ratio} or, as indicated above, the actual
number of contributing PUTs, those that have not been discarded due to
repeated sending by clients or (briefly) belonging to the previous
instance of the experiment. It reveals that, in some cases where the
amount of PUTs really increases, they are actually not contributing so
much, while light dots, corresponding to a ration close to 1, usually
stay in the lowest part of the graphic. 
%
\begin{figure}[ht!b]
\centering
<<entropy, cache=FALSE,echo=FALSE>>=
max <- max(new.experiment.data$IPs)
entropy.lm <- lm( new.experiment.data$milliseconds ~ new.experiment.data$entropy )
exp.palette <- colorRampPalette(brewer.pal(12,"Paired"))
ggplot(data=new.experiment.data)+geom_point(aes(x=entropy,y=milliseconds,size=PUTs,color=IPs))+scale_x_log10()+scale_y_log10()+scale_color_gradient(limits=c(1,max),low='red')
@ %def entropy data
%
\caption{Entropy of the user contribution vs. time to solution in the
  volunteer computing experiment. The fit to a linear model is significant. \label{fig:entropy}}
\end{figure}

This is interesting in the sense that in indicates that there might be
something more to performance that just the number of volunteers
contributing. In this experiment we include an additional measure:
user-contribution entropy, which is measure by computing the
compression entropy of the user contributions every for the whole
experiment.  This is computed converting the sequence of the number of
PUTs into a string and then compressing it using {\tt Zlib}, a
standard compression library. Since the compression rate is related to
the randomness of the sequence, a higher compression rate will
correspond to an experiment with a low entropy, while a lower
compression rate will indicate that there has been a certain amount of
unexpected changes during the experiment. Although less popular than
Jenssen-Shannon entropy, this way of measuring it is widely used
\cite{ornstein1993entropy} in case where we want to compute the
entropy of sequences of events, such as arrhythmias
\cite{baumert2004forecasting}. Unlike Jenssen-Shannon entropy, the
range of the compression entropy will vary between 0 (lower, or worse
compression rate) and 1 (no compression).

### 2016-11-14 EvoPar2017/paper.tex
Adds entropy linear model

In order to have enough data to perform more analysis on it, a new
experiment was performed along several days with, eventually, several
hundreds of persons participating on it. It was announced by several
popular Twitter figures and found its way into Reddit. The number of
PUTs per minute are shown in \ref{fig:minute}. Sudden peaks in them
are due to new announcements. Please check also the dot color, which
corresponds to the {\em ratio} or, as indicated above, the actual
number of contributing PUTs, those that have not been discarded due to
repeated sending by clients or (briefly) belonging to the previous
instance of the experiment. It reveals that, in some cases where the
amount of PUTs really increases, they are actually not contributing so
much, while light dots, corresponding to a ration close to 1, usually
stay in the lowest part of the graphic.

This is interesting in the sense that in indicates that there might be
something more to performance that just the number of volunteers
contributing. In this experiment we include an additional measure:
user-contribution entropy, which is measure by computing the
compression entropy of the user contributions every for the whole
experiment.  This is computed converting the sequence of the number of
PUTs into a string and then compressing it using {\tt Zlib}, a
standard compression library. Since the compression rate is related to
the randomness of the sequence, a higher compression rate will
correspond to an experiment with a low entropy, while a lower
compression rate will indicate that there has been a certain amount of
unexpected changes during the experiment. Although less popular than
Jenssen-Shannon entropy, this way of measuring it is widely used
\cite{ornstein1993entropy} in case where we want to compute the
entropy of sequences of events, such as arrhythmias
\cite{baumert2004forecasting}. Unlike Jenssen-Shannon entropy, the
range of the compression entropy will vary between 0 (lower, or worse
compression rate) and 1 (no compression). 
%
\begin{figure}[ht!b]
\centering
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
\includegraphics[width=\maxwidth]{figure/entropy-1}

@@

>>>>

In order to have enough data to perform more analysis on it, a new
experiment was performed along several days with, eventually, several
hundreds of persons participating on it. It was announced by several
popular Twitter figures and found its way into Reddit. The number of
PUTs per minute are shown in \ref{fig:minute}. Sudden peaks in them
are due to new announcements in social networks, usually; this is one
of the problems with this kind of volunteer computing systems: they
feed on the continuous influx of new users; however, this can be
automatized and also encouraged, and, as shown, once enough users have
participated it is sometimes easier to find new ones.

In Figure \ref{fig:minute} check
also the dot color, which 
corresponds to the {\em ratio} or, as indicated above, the actual
number of contributing PUTs, those that have not been discarded due to
repeated sending by clients or (briefly) belonging to the previous
instance of the experiment. It reveals that, in some cases where the
amount of PUTs really increases, they are actually not contributing so
much, while light dots, corresponding to a ration close to 1, usually
stay in the lowest part of the graphic. 
%
\begin{figure}[ht!b]
\centering
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
\includegraphics[width=\maxwidth]{figure/entropy-1}

@@

This is interesting in the sense that in indicates that there might be
something more to performance that just the number of volunteers
contributing. In this experiment we include an additional measure:
user-contribution entropy, which is measure by computing the
compression entropy of the user contributions every for the whole
experiment.  This is computed converting the sequence of the number of
PUTs into a string and then compressing it using {\tt Zlib}, a
standard compression library. Since the compression rate is related to
the randomness of the sequence, a higher compression rate will
correspond to an experiment with a low entropy, while a lower
compression rate will indicate that there has been a certain amount of
unexpected changes during the experiment. Although less popular than
Jenssen-Shannon entropy, this way of measuring it is widely used
\cite{ornstein1993entropy} in case where we want to compute the
entropy of sequences of events, such as arrhythmias
\cite{baumert2004forecasting}. Unlike Jenssen-Shannon entropy, the
range of the compression entropy will vary between 0 (lower, or worse
compression rate) and 1 (no compression).
